---
title: "fire_PA_clean"
author: "Emma Rieves & Zac Clement"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages
```{r}
library(plyr)
library(tidyverse)
library(stringr)
library(readr)
library(gtsummary)
library(sf)
library(kableExtra)
```


Cleaning/correction information: https://fire.airnow.gov/#correction-equation

# Load data
```{r}
# read in the file
AQ_df = read_csv("../intermediary_outputs/aq_data.csv")

# drop the index columns
AQ_df = AQ_df[,-1]

```


# Cleaning
For 10-min averaged data

    * The data point is valid if the A and B channel PM2.5 measurements are within either 5 micrograms per cubic meter or 70% relative percent difference
    * The A and B channels are averaged
```{r}
# determine if data points are valid based on EPA logic
AQ_df = AQ_df %>%
  rowwise() %>% 
  mutate(
    a_b_agree = 
      # if eiher pm25_a or pm25_b is NA, disagree
      ifelse(is.na(pm25_a) | is.na(pm25_b), "disagree",
          # if pm25_a is less than pm25_b + 5
          ifelse(pm25_a <= pm25_b + 5,"agree",
                 # if pm25_a is more than pm25_b - 5
                 ifelse(pm25_a >= pm25_b - 5, "agree",
                        # if pm25_a is within 70% relative difference
                        ifelse(
                          abs(((pm25_a - pm25_b)/pm25_b)*100) >= 70 | abs(((pm25_b - pm25_a)/pm25_a)*100) >= 70, "agree","disagree")))))


# select only valid points, average a and b channels for those data
(AQ_df_valid = AQ_df %>% 
  dplyr::filter(a_b_agree == "agree") %>%
  rowwise() %>% 
  mutate(pm_avg = mean(c(pm25_a,pm25_b)),
         ab_difference = pm25_a-pm25_b))
```

## Cleaning summary stats

```{r}
agree = AQ_df %>%
  dplyr::filter(a_b_agree == "agree") %>%
  group_by(ID, a_b_agree) %>%
  dplyr::summarise(n_agree = n())

disagree = AQ_df %>%
  dplyr::filter(a_b_agree != "agree") %>%
  group_by(ID, a_b_agree) %>%
  dplyr::summarise(n_disagree = n())
```

```{r}
agree_counts = inner_join(agree, disagree, by="ID") %>%
  dplyr::select(-c(a_b_agree.x, a_b_agree.y)) %>%
  mutate(pct_agree = 100 * n_agree / (n_agree + n_disagree),
         pct_disagree = 100 * n_disagree / (n_agree + n_disagree)) %>%
  arrange(desc(pct_disagree))

agree_counts %>%
  kbl(digits=2, caption="Number of Datapoints that Agree Between Channels, Per Sensor") %>%
  kable_styling()
```

```{r}
# how many sensors have significant % disagreement btwn a & b?
agree_counts = agree_counts %>%
  mutate(group = case_when(pct_disagree >= 90 ~ 'very high (>90%)',
                           pct_disagree < 90 & pct_disagree >= 75 ~ 'high (75%-90%)',
                           pct_disagree < 75 & pct_disagree >= 50 ~ 'moderate (50%-75%)',
                           pct_disagree < 50 & pct_disagree >= 25 ~ 'low (25%-50%)',
                           pct_disagree < 25 & pct_disagree >= 1 ~ 'very low (1%-25%)',
                           pct_disagree < 1 ~ 'less than 1%')) %>%
  mutate(group = factor(group, levels=c("very high (>90%)", "high (75%-90%)", "moderate (50%-75%)", 'low (25%-50%)', 'very low (1%-25%)', 'less than 1%')))

agree_counts %>% group_by(group) %>%
  dplyr::summarise(n = n()) %>%
  kbl(caption="Number of Sensors with Their Two Channels Disagreeing") %>%
  kable_styling()
```

```{r}
100 * sum(agree_counts$n_disagree) / (sum(agree_counts$n_disagree) + sum(agree_counts$n_agree))
```
5.80% of points are lost by using the EPA formula.

## Mapping
```{r}
# BOUNDARIES
# Boulder boundaries
BO_CO = st_read("../GIS_inputs_destruction_fireboundary/Boulder_county_munis/Municipalities.shp")
fire_counties = BO_CO %>%
  dplyr::filter(ZONEDESC == "Louisville" |
                  ZONEDESC == "Superior" |
                  ZONEDESC == "Broomfield" |
                  ZONEDESC == "Lafayette" |
                  ZONEDESC == "Boulder")
boulder_precincts = st_read("../GIS_inputs_destruction_fireboundary/Unincorporated_Boulder/Unincorporated_Boulder.shp")
broomfield_precincts = st_read("../GIS_inputs_destruction_fireboundary/Broomfield_Precincts/Precincts.shp")
westminster_city = st_read("../GIS_inputs_destruction_fireboundary/Westminster_CityLimits/CityLimits.shp")

# save proj string for fire-affected counties to use for other transformations
prg = raster::crs(fire_counties,asText=TRUE)

# Marshall fire boundary
wfigs_fire = st_read("../GIS_inputs_destruction_fireboundary/WFIGS_-_Wildland_Fire_Perimeters_Full_History/FH_Perimeter.shp")

# filter fire to marshall fire only; update crs
marshall_fire = wfigs_fire %>% filter(poly_Incid == "Marshall") %>% st_transform(crs = st_crs(prg))

## sensor info
sensors = read.csv("../intermediary_outputs/sensor_data_full.csv")
sensors = st_as_sf(sensors, coords=c("Lon", "Lat"), crs=st_crs(prg))
```

```{r}
ggplot(boulder_precincts) +
  geom_sf() +
  geom_sf(data=broomfield_precincts, color="darkblue") +
  geom_sf(data=westminster_city, color="darkgreen") +
  geom_sf(data=marshall_fire$geometry, fill="red", alpha=0.5) +
  geom_sf(data=sensors$geometry) +
  ggtitle("All Sensor Locations")
```

```{r}
# join the sensor locations to the agree/disagree information
agree_counts = left_join(agree_counts, sensors, by="ID")
```


```{r}
ggplot(boulder_precincts) +
  geom_sf(color="gray") +
  geom_sf(data=broomfield_precincts, color="gray") +
  geom_sf(data=westminster_city, color="gray") +
  geom_sf(data=marshall_fire$geometry, fill="red", alpha=0.5) +
  geom_sf(data=agree_counts, aes(col=pct_disagree, geometry=geometry)) +
  scale_color_viridis_c() +
  ggtitle("All Sensor Locations 12/30/21 - 05/01/22")
```

## Limiting to fire period
```{r}
agree = AQ_df %>%
  dplyr::filter(a_b_agree == "agree" & datetime < "2022-01-02") %>%
  group_by(ID, a_b_agree) %>%
  dplyr::summarise(n_agree = n())

disagree = AQ_df %>%
  dplyr::filter(a_b_agree != "agree" & datetime < "2022-01-02") %>%
  group_by(ID, a_b_agree) %>%
  dplyr::summarise(n_disagree = n())

agree_counts_fire = inner_join(agree, disagree, by="ID") %>%
  dplyr::select(-c(a_b_agree.x, a_b_agree.y)) %>%
  mutate(pct_agree = 100 * n_agree / (n_agree + n_disagree),
         pct_disagree = 100 * n_disagree / (n_agree + n_disagree)) %>%
  arrange(desc(pct_disagree))

agree_counts_fire %>%
  kbl(digits=2, caption="Number of Datapoints that Agree Between Channels, Per Sensor") %>%
  kable_styling()
```

```{r}
agree_counts_fire = left_join(agree_counts_fire, sensors, by="ID")
```

```{r}
ggplot(boulder_precincts) +
  geom_sf(color="gray") +
  geom_sf(data=broomfield_precincts, color="gray") +
  geom_sf(data=westminster_city, color="gray") +
  geom_sf(data=marshall_fire$geometry, fill="red", alpha=0.5) +
  geom_sf(data=agree_counts_fire, aes(col=pct_disagree, geometry=geometry), alpha=0.75) +
  scale_color_viridis_c() +
  ggtitle("All Sensor Locations 12/30/21 - 01/02/22")
```

Only looking at disagreeing sensors
```{r}
only_disagree <- agree_counts_fire %>%
  filter(pct_disagree > 0)

ggplot(boulder_precincts) +
  geom_sf(color="gray") +
  geom_sf(data=broomfield_precincts, color="gray") +
  geom_sf(data=westminster_city, color="gray") +
  geom_sf(data=marshall_fire$geometry, fill="red", alpha=0.5) +
  geom_sf(data=only_disagree, aes(col=pct_disagree, geometry=geometry, alpha = 0.5, size=pct_disagree)) +
  scale_color_viridis_c() +
  ggtitle("All Sensor Locations 12/30/21 - 01/02/22")
```

```{r}
ggplot(data=agree_counts_fire) +
  geom_histogram(aes(x=pct_disagree), fill="black", color="white") +
  ggtitle("Histogram of the percentage of data points in fire period that disagree")
```

All of the data points that disagree are thrown out with the EPA cleaning & correcting. This means we lose a lot of the actual fire data (see map above).
```{r}
sum(agree_counts_fire$n_disagree)
sum(agree_counts_fire$n_disagree) + sum(agree_counts_fire$n_agree)

100 * sum(agree_counts_fire$n_disagree) / (sum(agree_counts_fire$n_disagree) + sum(agree_counts_fire$n_agree))
```

29.88% of the data between 12/30/21 and 01/02/22 is lost when we clean via the EPA standards.

## Same Analysis, but January
```{r}
agree = AQ_df %>%
  dplyr::filter(a_b_agree == "agree" & datetime < "2022-02-01") %>%
  group_by(ID, a_b_agree) %>%
  dplyr::summarise(n_agree = n())

disagree = AQ_df %>%
  dplyr::filter(a_b_agree != "agree" & datetime < "2022-02-01") %>%
  group_by(ID, a_b_agree) %>%
  dplyr::summarise(n_disagree = n())

agree_counts_jan = inner_join(agree, disagree, by="ID") %>%
  dplyr::select(-c(a_b_agree.x, a_b_agree.y)) %>%
  mutate(pct_agree = 100 * n_agree / (n_agree + n_disagree),
         pct_disagree = 100 * n_disagree / (n_agree + n_disagree)) %>%
  arrange(desc(pct_disagree))

agree_counts_jan %>%
  kbl(digits=2, caption="Number of Datapoints that Agree Between Channels, Per Sensor") %>%
  kable_styling()
```

```{r}
agree_counts_jan = left_join(agree_counts_jan, sensors, by="ID")
```

```{r}
ggplot(boulder_precincts) +
  geom_sf(color="gray") +
  geom_sf(data=broomfield_precincts, color="gray") +
  geom_sf(data=westminster_city, color="gray") +
  geom_sf(data=marshall_fire$geometry, fill="red", alpha=0.5) +
  geom_sf(data=agree_counts_jan, aes(col=pct_disagree, geometry=geometry)) +
  scale_color_viridis_c() +
  ggtitle("All Sensor Locations 01/01/22 - 02/01/22")
```

Only looking at disagreeing sensors
```{r}
only_disagree_jan <- agree_counts_jan %>%
  filter(pct_disagree > 0)

ggplot(boulder_precincts) +
  geom_sf(color="gray") +
  geom_sf(data=broomfield_precincts, color="gray") +
  geom_sf(data=westminster_city, color="gray") +
  geom_sf(data=marshall_fire$geometry, fill="red", alpha=0.5) +
  geom_sf(data=only_disagree_jan, aes(col=pct_disagree, geometry=geometry, alpha = 0.5, size=pct_disagree)) +
  scale_color_viridis_c() +
  ggtitle("All Sensor Locations 12/30/21 - 01/02/22")
```

```{r}
sum(agree_counts_jan$n_disagree)
sum(agree_counts_jan$n_disagree) + sum(agree_counts_jan$n_agree)

100 * sum(agree_counts_jan$n_disagree) / (sum(agree_counts_jan$n_disagree) + sum(agree_counts_jan$n_agree))
```

8.58% of the data between 01/01/22 and 02/01/22 is lost when we clean via the EPA standards.

# Aggregated Agreement by Day/Time
## January
```{r}
jan_agg <- AQ_df %>%
  dplyr::filter(datetime < "2022-02-01") %>%
  group_by(datetime) %>%
  dplyr::summarise(total = n(),
                   num_disagree = sum(a_b_agree == "disagree"),
                   num_agree = sum(a_b_agree == "aggree")) %>%
  mutate(pct_disagree = num_disagree / total)
```

```{r}
ggplot(jan_agg, aes(x=datetime, y=pct_disagree)) +
  geom_line() +
  geom_smooth()
```

### Aggregated by date
```{r}
jan_agg1 <- jan_agg %>%
  mutate(date = as.Date(datetime),
         pct_disagree = num_disagree / total) %>%
  group_by(date) %>%
  dplyr::summarise(pct_disagree = mean(pct_disagree))

ggplot(jan_agg1, aes(x=date, y=pct_disagree)) +
  geom_point() +
  geom_line() +
  geom_smooth()
```


## Fire period
```{r}
fire_agg <- AQ_df %>%
  dplyr::filter(datetime < "2022-01-02") %>%
  group_by(datetime) %>%
  dplyr::summarise(total = n(),
                   num_disagree = sum(a_b_agree == "disagree"),
                   num_agree = sum(a_b_agree == "aggree")) %>%
  mutate(pct_disagree = num_disagree / total)
```

```{r}
# 11 am on 12/31/2021 was the first 911 call about the fire
ggplot(fire_agg, aes(x=datetime, y=pct_disagree)) +
  geom_point() +
  geom_smooth() +
  geom_vline(xintercept=as.POSIXct("2021-12-31 11:00:00", tz="UTC"), color="red") +
  ggtitle("Percent of Disagreeing Data for the Fire Period (line is at 2021-12-31 11:00:00)")
```
There's high disagreement right before the fire was reported -> purple air sensors heavily affected by wind?

# Summary Stats
```{r}
# summary stats -- pct NA for each channel
AQ_df %>% group_by(ID) %>% summarise(count_A_na = sum(is.na(pm25_a)),
                                     count_B_na = sum(is.na(pm25_b)),
                                     pct_A_na = (count_A_na/length(pm25_a))*100,
                                     pct_B_na = (count_B_na/length(pm25_b))*100)

# overall percent valid
AQ_df %>% group_by(ID) %>% summarise(pct_valid = sum(a_b_agree == "agree")/length(pm25_a) * 100)

# average difference between A and B channels for each sensor
# typically good agreement between the A and B channels..
AQ_df_valid %>% group_by(ID) %>% dplyr::summarise(ab_diff_summary = format(mean(ab_difference),scientific=FALSE))
```

```{r}
## work on the cross table -- maybe add in time period and municipality as organizing vars
AQ_df %>% tbl_cross(row = ID, col = a_b_agree, percent = "row") %>% as_gt()
```



# Correction
equation found here: https://fire.airnow.gov/#correction-equation

    * Low Concentration
    PAcf_atm < 50 µg/m3 	
    PM2.5 = 0.52 x (PAcf_atm) - 0.086 x RH + 5.75
    
    * Mid Concentration 
    50 µg/m3 > (PAcf_atm) <229 	
    PM2.5 = 0.786 x (PAcf_atm) - 0.086 x RH + 5.75
    
    * High Concentration
    PAcf_atm > 229 µg/m3 	
    PM2.5 = 0.69 x (PAcf_atm) + 8.84 x 10^(-4) x PAcf_atm^2 + 2.97
    
    
```{r}
# linear piecewise correction from the EPA
## question about using the pm avg or just the a channel
## confirm that this is the best correction
(corrected_AQ = AQ_df_valid %>% 
   rowwise() %>% 
   mutate(
  corrected_pm = ifelse(pm_avg < 50, ((0.52 * pm_avg) - (0.088 * rh) + 5.75),
                        ifelse(pm_avg >= 50 | pm_avg < 299, ((0.786 * pm_avg) - (0.086 * rh) + 5.75),
                               ifelse(pm_avg >= 229, ((0.69 * pm_avg) + 8.84 * (0.001 * pm_avg^2) + 2.97), NA)))))
```

```{r}
# STATS
# summary stats -- AB difference (for corrected channels)
## this could be helpful in determining whether the step of removing observations in cleaning is important
summary(corrected_AQ$ab_difference)

# mean difference between A and B channels per sensor
(mean_sensor_AB_diff = corrected_AQ %>% group_by(ID) %>% dplyr::summarise(mean_ab_diff = mean(ab_difference)))

# distribution of differences between sensors -- indicates that some sensor differenes are > 3000 on average! will investigate further
hist(mean_sensor_AB_diff$mean_ab_diff)

# create a new version exploring AB differences, but excluding outliers
mean_sensor_AB_diff_sub1000 = mean_sensor_AB_diff %>% filter(mean_ab_diff < 1000)

# most AB differences are small
hist(mean_sensor_AB_diff_sub1000$mean_ab_diff,breaks = 20)
```

```{r}
# STATS
# summary stats -- corrected pm
# weird that this contains negatives.. the EPA info doesn't mention anything about that
summary(corrected_AQ$corrected_pm)

# mean corrected PM
(mean_AQ = corrected_AQ %>% group_by(ID) %>% dplyr::summarise(mean_pm = mean(corrected_pm)))

# distribution of differences between sensors -- indicates that some sensor differecnes are > 3000 on average! will investigate further
hist(mean_sensor_AB_diff$mean_ab_diff)

# create a new version exploring AB differences, but excluding outliers
mean_sensor_AB_diff_sub1000 = mean_sensor_AB_diff %>% filter(mean_ab_diff < 1000)

# most AB differences are small
hist(mean_sensor_AB_diff_sub1000$mean_ab_diff,breaks = 20)
```



# Export data
```{r}
# AQ df for cleaning
write.csv(corrected_AQ,"../intermediary_outputs/corrected_AQ_data.csv")
```

